A Bioconductor workflow for processing and analysing spatial proteomics data

### 2018-04-11 bioc-workflow-resubmit.tex
verison for resubmission

While we focus here on a LOPIT-type dataset, these analyses are
\hl{relevent} for any quantitative spatial proteomics data, irrespective of
the fractionation or quantitation (i.e. labelled or label-free)
methods.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{combined} \hlkwb{<-} \hlkwd{combine}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
\end{alltt}

{\ttfamily\noindent\color{warningcolor}{\#\# Warning: 'combine' is deprecated.\\\#\# Use 'gtable\_combine' instead.\\\#\# See help("{}Deprecated"{})}}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in `[[<-.data.frame`(`*tmp*`, i, value = structure(list(t = numeric(0), : replacement has 0 rows, data has 10}}\begin{alltt}
\hlstd{combined}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 5032 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
##     total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
##   pubMedIds: 26754106 
## Annotation:  
## - - - Processing information - - -
## Combined [5489,10] and [6268,10] MSnSets Tue Apr 10 11:36:36 2018 
## Subset [6725,20][5032,20] Tue Apr 10 11:36:36 2018 
## Removed features with more than 0 NAs: Tue Apr 10 11:36:36 2018 
## Dropped featureData's levels Tue Apr 10 11:36:36 2018 
## Subset [5032,20][5032,20] Tue Apr 10 14:00:41 2018 
## Removed features with more than 0 NAs: Tue Apr 10 14:00:41 2018 
## Dropped featureData's levels Tue Apr 10 14:00:41 2018 
## Subset [5032,20][5032,20] Tue Apr 10 16:11:22 2018 
## Removed features with more than 0 NAs: Tue Apr 10 16:11:22 2018 
## Dropped featureData's levels Tue Apr 10 16:11:22 2018 
## Subset [5032,20][5032,20] Tue Apr 10 16:17:52 2018 
## Removed features with more than 0 NAs: Tue Apr 10 16:17:52 2018 
## Dropped featureData's levels Tue Apr 10 16:17:52 2018 
## Subset [5032,20][5032,20] Tue Apr 10 16:25:51 2018 
## Removed features with more than 0 NAs: Tue Apr 10 16:25:51 2018 
## Dropped featureData's levels Tue Apr 10 16:25:51 2018 
## Subset [5032,20][5032,20] Tue Apr 10 16:29:26 2018 
## Removed features with more than 0 NAs: Tue Apr 10 16:29:26 2018 
## Dropped featureData's levels Tue Apr 10 16:29:27 2018 
## Subset [5032,20][5032,20] Tue Apr 10 16:35:47 2018 
## Removed features with more than 0 NAs: Tue Apr 10 16:35:47 2018 
## Dropped featureData's levels Tue Apr 10 16:35:47 2018 
## Subset [5032,20][5032,20] Tue Apr 10 16:47:20 2018 
## Removed features with more than 0 NAs: Tue Apr 10 16:47:20 2018 
## Dropped featureData's levels Tue Apr 10 16:47:20 2018 
##  MSnbase version: 2.5.9
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\end{knitrout}
\caption{Heatmap of missing values. Note that the features are
  re-ordered to highlight \hl{clusters} of proteins with similar numbers of
  missing values.}
  \label{fig:namap}
\end{figure}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{combined} \hlkwb{<-} \hlkwd{filterNA}\hlstd{(combined)}
\hlstd{combined}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 5032 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
##     total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
##   pubMedIds: 26754106 
## Annotation:  
## - - - Processing information - - -
## Combined [5489,10] and [6268,10] MSnSets Tue Apr 10 11:36:36 2018 
## Subset [6725,20][5032,20] Tue Apr 10 11:36:36 2018 
## Removed features with more than 0 NAs: Tue Apr 10 11:36:36 2018 
## Dropped featureData's levels Tue Apr 10 11:36:36 2018 
## Subset [5032,20][5032,20] Tue Apr 10 14:00:41 2018 
## Removed features with more than 0 NAs: Tue Apr 10 14:00:41 2018 
## Dropped featureData's levels Tue Apr 10 14:00:41 2018 
## Subset [5032,20][5032,20] Tue Apr 10 16:11:22 2018 
## Removed features with more than 0 NAs: Tue Apr 10 16:11:22 2018 
## Dropped featureData's levels Tue Apr 10 16:11:22 2018 
## Subset [5032,20][5032,20] Tue Apr 10 16:17:52 2018 
## Removed features with more than 0 NAs: Tue Apr 10 16:17:52 2018 
## Dropped featureData's levels Tue Apr 10 16:17:52 2018 
## Subset [5032,20][5032,20] Tue Apr 10 16:25:51 2018 
## Removed features with more than 0 NAs: Tue Apr 10 16:25:51 2018 
## Dropped featureData's levels Tue Apr 10 16:25:51 2018 
## Subset [5032,20][5032,20] Tue Apr 10 16:29:26 2018 
## Removed features with more than 0 NAs: Tue Apr 10 16:29:26 2018 
## Dropped featureData's levels Tue Apr 10 16:29:27 2018 
## Subset [5032,20][5032,20] Tue Apr 10 16:35:47 2018 
## Removed features with more than 0 NAs: Tue Apr 10 16:35:47 2018 
## Dropped featureData's levels Tue Apr 10 16:35:47 2018 
## Subset [5032,20][5032,20] Tue Apr 10 16:47:20 2018 
## Removed features with more than 0 NAs: Tue Apr 10 16:47:20 2018 
## Dropped featureData's levels Tue Apr 10 16:47:20 2018 
## Subset [5032,20][5032,20] Tue Apr 10 16:52:54 2018 
## Removed features with more than 0 NAs: Tue Apr 10 16:52:54 2018 
## Dropped featureData's levels Tue Apr 10 16:52:54 2018 
##  MSnbase version: 2.5.9
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\hl{This is further highlighted if we plot the
profiles of these two clusters using the \texttt{plotDist} function}
(Figure \ref{fig:plotDist2}). \hl{The \texttt{plotDist} function is
another useful visualisation that relies on marker annotation. It
allows one to represent the protein profiles occupancy along the
gradient. While the PCA
plot enables efficient visualisation of the complete dataset and
assessment the relative separation of different sub-cellular niches,
comparing profiles of a few marker clusters is useful to assess how
exactly they differ (in terms of peak channels, for example). On
figure} \ref{fig:plotDist2}, \hl{we plot the profile of the mitochondrial
and peroxisome markers to highlight the differences in profiles
between these two sets of markers in the channels labelled with tag
129C, as represented above along the 7th PC on the PCA plot on
figure} \ref{fig:plotmarkers}.

@@

\subsection*{Features of interest}
In addition to adding annotation using the \texttt{addMarkers}
function, one can store specific sets of proteins by using the
\textit{Features of interest} infrastructure from the
\Biocpkg{MSnbase} package. If users have specific subsets of proteins
they wish to highlight in their data (possibly across multiple
experiments) they would first create a \texttt{FeaturesOfInterest}
object \hl{and then use the \textit{highlightOnPlot} function to 
visualise these}. For example, if we wanted to highlight \st{a} 
proteins with the accession numbers Q8CG48, Q8CG47, Q8K2Z4, and Q8C156, 
which are some of the proteins known to form part of the 13S condensin complex, 
we \hl{would call the following code:}

\begin{figure}[!ht]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{prots} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"Q8CG48"}\hlstd{,} \hlstr{"Q8CG47"}\hlstd{,} \hlstr{"Q8K2Z4"}\hlstd{,} \hlstr{"Q8C156"}\hlstd{)}
\hlstd{foi13s} \hlkwb{<-} \hlkwd{FeaturesOfInterest}\hlstd{(}\hlkwc{description} \hlstd{=} \hlstr{"13S consensin proteins"}\hlstd{,}
                             \hlkwc{fnames} \hlstd{= prots,}
                             \hlkwc{object} \hlstd{= hl)}
\hlstd{foi13s}
\end{alltt}
\begin{verbatim}
## Traceable object of class "FeaturesOfInterest"
##  Created on Tue Apr 10 16:52:56 2018 
##  Description:
##   13S consensin proteins
##  4 features of interest:
##    Q8CG48, Q8CG47, Q8K2Z4, Q8C156
\end{verbatim}
\begin{alltt}
\hlkwd{plot2D}\hlstd{(hl)}
\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{cex} \hlstd{=} \hlnum{.6}\hlstd{)}
\hlkwd{highlightOnPlot}\hlstd{(hl, foi13s)}
\end{alltt}
\end{kframe}
\includegraphics[width=.8\textwidth]{figure/foi-1}

@@

Direct comparisons of individual channels in replicated experiments
do\st{es} not provide an adequate, goal-driven assessment of different
experiments. Indeed, due to the nature of the experiment and gradient
fraction collection, the quantitative channels do not correspond to
identical selected fractions along the gradient. For example, in the
table below (taken from \texttt{hl}'s \texttt{pData}) TMT channels
127C (among others) in both replicates originate from different sets
of gradient fractions (gradient fractions 7 - 9 and 8 - 9 for each
replicate, respectively). Different sets of gradient fractions are
often pooled to obtain enough material and optimise acurate
quantitation.

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Tue Apr 10 16:52:56 2018
\begin{table}[ht]
\centering
\begin{tabular}{rrllr}
  \hline
 & Replicate & Tag & Gradient.Fraction & Iodixonal.Density \\ 
  \hline
X127C &   1 & 127C & 8 to 9 (pooled) & 11.00 \\ 
  X127C.1 &   2 & 127C & 7 to 9 (pooled) & 10.00 \\ 
   \hline
\end{tabular}
\caption{Differences in gradient fraction pooling.} 
\label{tab:pdtab}
\end{table}

The more \hl{relevent} comparison unit is not a single channel, but rather
the complete protein occupancy profiles, which are best visualised 
experiment-wide on a PCA plot. As such, we prefer to focus on the
direct, qualitative comparison of individual replicate PCA plots,
assuring that each displays acceptable sub-cellular resolution. Note
that in the code chunk below, we mirror the x-axis to represent the
two figures with the same orientation. The interactive "compare"
application part of the \Biocpkg{pRolocGUI} package is also useful for
examining replicate experiments (see the next section interactive
visualisation for details).

@@

\hl{The above analysis is comptationally intensive and best parallelised over
multiple workers. This phenoDisco analysis took ~24 hours to complete
when parallelised over 40 workers. As such, to avoid executing this
time consuming analysis users can access the above results which are
pre-computed and stored along with the \texttt{pRolocdata} package. 
Please see the appendix to load these results.}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{hl}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 5032 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126 X127N ... X131.1 (20 total)
##   varLabels: Replicate Tag Gradient.Fraction Iodixonal.Density
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
##   fvarLabels: uniprot.accession uniprot.id ... pd (27 total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Added markers from  'mrk' marker vector. Tue Apr 10 16:52:55 2018 
## Added markers from  'pdres' marker vector. Tue Apr 10 16:52:56 2018 
##  MSnbase version: 2.5.10
\end{verbatim}
\begin{alltt}
\hlkwd{getMarkers}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"pd"}\hlstd{)}
\end{alltt}
\begin{verbatim}
## organelleMarkers
##                          40S Ribosome 
##                                   106 
##                          60S Ribosome 
##                                    95 
## Endoplasmic reticulum/Golgi apparatus 
##                                   393 
##                         Mitochondrion 
##                                   525 
##                           Phenotype 1 
##                                   300 
##                           Phenotype 2 
##                                   253 
##                           Phenotype 3 
##                                   203 
##                           Phenotype 4 
##                                    74 
##                           Phenotype 5 
##                                    91 
##                       Plasma membrane 
##                                   421 
##                            Proteasome 
##                                    92 
##                               unknown 
##                                  2479
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\hl{In the code chunk below we then pass these weights 
to the \texttt{svmOptimisation} function. Once again, we provide
the optimisation results for users to load directly if they wish to save
computational time, please see the appendix to load these results.}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{svmf} \hlkwb{<-} \hlkwd{dir}\hlstd{(extdatadir,} \hlkwc{full.names} \hlstd{=} \hlnum{TRUE}\hlstd{,}
              \hlkwc{pattern} \hlstd{=} \hlstr{"bpw-svmopt.rds"}\hlstd{)}
\hlstd{params} \hlkwb{<-} \hlkwd{readRDS}\hlstd{(svmf)}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{hl} \hlkwb{<-} \hlkwd{svmClassification}\hlstd{(hl, params,} \hlkwc{class.weights} \hlstd{= w,} \hlkwc{fcol} \hlstd{=} \hlstr{"markers"}\hlstd{)}
\end{alltt}

{\ttfamily\noindent\color{warningcolor}{\#\# Warning in seq\_len(trpr): first element used of 'length.out' argument}}\end{kframe}
\end{knitrout}

@@

There are many ways to set thresholds and the choice of method will
depend on the biological question and experimental design at hand. One
viable approach in the frame of the above experimetal design would be
to manually set a FDR, say 5\%, per organelle. To do this the user
would examine the top scoring predictions for each organelle, and then
set a threshold at the score at which they achieve 5\% of false
assignments per organelle.  The definintion of a false assignment
would depend on the information available, for example, validity or
lack of validity for the localisation from another experiment as
reported in the literature or a reliable database.  If such
information is not available, one crude method is to set a threshold
per organelle by extracting the median or 3rd \hl{quartile} score per
organelle.  For example, in the code chunk below, we use the
\texttt{orgQuants} function to extract the median organelle scores and
then pass these scores to the \texttt{getPredictions} function to
extract the new localisations that meet this scoring criteria. Any
sub-cellular predictions that fall below the specified thresholds are
labelled as unknown.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{(ts} \hlkwb{<-} \hlkwd{orgQuants}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{=} \hlnum{.5}\hlstd{))}
\end{alltt}
\begin{verbatim}
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##               0.4323379               0.3037630               0.3822328 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##               0.6930642               0.6014878               0.4426545 
##    Extracellular matrix                Lysosome           Mitochondrion 
##               0.4224179               0.5978957               0.9497788 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##               0.7950203               0.7117761               0.3147679 
##         Plasma membrane              Proteasome 
##               0.7042806               0.4150910
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##               0.4323379               0.3037630               0.3822328 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##               0.6930642               0.6014878               0.4426545 
##    Extracellular matrix                Lysosome           Mitochondrion 
##               0.4224179               0.5978957               0.9497788 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##               0.7950203               0.7117761               0.3147679 
##         Plasma membrane              Proteasome 
##               0.7042806               0.4150910
\end{verbatim}
\begin{alltt}
\hlstd{hl} \hlkwb{<-} \hlkwd{getPredictions}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{= ts)}
\end{alltt}
\begin{verbatim}
## ans
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##                      85                     170                      84 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##                     301                     478                      99 
##    Extracellular matrix                Lysosome           Mitochondrion 
##                      27                     123                     523 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##                     230                     341                      38 
##         Plasma membrane              Proteasome                 unknown 
##                     322                     154                    2057
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

We generally find supervised learning more suited to the task of
protein localisation prediction in which we use high-quality curated
marker proteins to build a classifier, instead of using an entirely
unsupervised approach to look for clusters and then look for
enrichment of organelles and complexes. In the latter we do not make
good use of valuable prior knowledge, and in our experience
unsupervised clustering can be extremely difficult due to poor
estimates of the number of clusters that may appear in the data.

@@

##  [1] gridExtra_2.3        pRolocdata_1.17.3    pRoloc_1.19.3       

>>>>

While we focus here on a LOPIT-type dataset, these analyses are
relevent for any quantitative spatial proteomics data,  
irrespective of the fractionation (i.e. density gradient or 
differential centrifugation \cite{Itzhak:2016}) or quantitation 
(i.e. labelled or label-free) methods.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{combined} \hlkwb{<-} \hlstd{MSnbase}\hlopt{::}\hlkwd{combine}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
\hlstd{combined}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 6725 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2Y3-3 (6725 total)
##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
##     total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
##   pubMedIds: 26754106 
## Annotation:  
## - - - Processing information - - -
## Combined [5489,10] and [6268,10] MSnSets Wed Apr 11 14:12:49 2018 
##  MSnbase version: 2.5.9
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\end{knitrout}
\caption{Heatmap of missing values. Note that the features are
  re-ordered to highlight clusters of proteins with similar numbers of
  missing values.}
  \label{fig:namap}
\end{figure}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{combined} \hlkwb{<-} \hlkwd{filterNA}\hlstd{(combined)}
\hlstd{combined}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 5032 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
##     total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
##   pubMedIds: 26754106 
## Annotation:  
## - - - Processing information - - -
## Combined [5489,10] and [6268,10] MSnSets Wed Apr 11 14:12:49 2018 
## Subset [6725,20][5032,20] Wed Apr 11 14:12:50 2018 
## Removed features with more than 0 NAs: Wed Apr 11 14:12:50 2018 
## Dropped featureData's levels Wed Apr 11 14:12:50 2018 
##  MSnbase version: 2.5.9
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

This is further highlighted if we plot the
profiles of these two clusters using the \texttt{plotDist} function
(Figure \ref{fig:plotDist2}). The \texttt{plotDist} function is
another useful visualisation that relies on marker annotation. It
allows one to represent the protein profiles occupancy along the
gradient. While the PCA
plot enables efficient visualisation of the complete dataset and
assessment the relative separation of different sub-cellular niches,
comparing profiles of a few marker clusters is useful to assess how
exactly they differ (in terms of peak channels, for example). On
figure \ref{fig:plotDist2}, we plot the profiles of the 
mitochondrial and peroxisome markers to highlight the differences 
in the channels labelled with tag 129C, also represented above  
along the 7th PC on the PCA plot on figure \ref{fig:plotmarkers}.

@@

\subsection*{Features of interest}
In addition to adding annotation using the \texttt{addMarkers}
function, one can store specific sets of proteins by using the
\textit{Features of interest} infrastructure from the
\Biocpkg{MSnbase} package. If users have specific subsets of proteins
they wish to highlight in their data (possibly across multiple
experiments) they would first create a \texttt{FeaturesOfInterest}
object and then use the \textit{highlightOnPlot} function to 
visualise these. For example, if we wanted to highlight 
proteins with the accession numbers Q8CG48, Q8CG47, Q8K2Z4, and 
Q8C156, which are some of the proteins known to form part of the 
13S condensin complex, we would call the following code:

\begin{figure}[!ht]
  \centering
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{prots} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"Q8CG48"}\hlstd{,} \hlstr{"Q8CG47"}\hlstd{,} \hlstr{"Q8K2Z4"}\hlstd{,} \hlstr{"Q8C156"}\hlstd{)}
\hlstd{foi13s} \hlkwb{<-} \hlkwd{FeaturesOfInterest}\hlstd{(}\hlkwc{description} \hlstd{=} \hlstr{"13S consensin proteins"}\hlstd{,}
                             \hlkwc{fnames} \hlstd{= prots,}
                             \hlkwc{object} \hlstd{= hl)}
\hlstd{foi13s}
\end{alltt}
\begin{verbatim}
## Traceable object of class "FeaturesOfInterest"
##  Created on Wed Apr 11 14:12:52 2018 
##  Description:
##   13S consensin proteins
##  4 features of interest:
##    Q8CG48, Q8CG47, Q8K2Z4, Q8C156
\end{verbatim}
\begin{alltt}
\hlkwd{plot2D}\hlstd{(hl)}
\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{cex} \hlstd{=} \hlnum{.6}\hlstd{)}
\hlkwd{highlightOnPlot}\hlstd{(hl, foi13s)}
\end{alltt}
\end{kframe}
\includegraphics[width=.8\textwidth]{figure/foi-1}

@@

Direct comparisons of individual channels in replicated experiments
do not provide an adequate, goal-driven assessment of different
experiments. Indeed, due to the nature of the experiment and gradient
fraction collection, the quantitative channels do not correspond to
identical selected fractions along the gradient. For example, in the
table below (taken from \texttt{hl}'s \texttt{pData}) TMT channels
127C (among others) in both replicates originate from different sets
of gradient fractions (gradient fractions 7 - 9 and 8 - 9 for each
replicate, respectively). Different sets of gradient fractions are
often pooled to obtain enough material and optimise acurate
quantitation.

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Wed Apr 11 14:12:52 2018
\begin{table}[ht]
\centering
\begin{tabular}{rrllr}
  \hline
 & Replicate & Tag & Gradient.Fraction & Iodixonal.Density \\ 
  \hline
X127C &   1 & 127C & 8 to 9 (pooled) & 11.00 \\ 
  X127C.1 &   2 & 127C & 7 to 9 (pooled) & 10.00 \\ 
   \hline
\end{tabular}
\caption{Differences in gradient fraction pooling.} 
\label{tab:pdtab}
\end{table}

The more relevent comparison unit is not a single channel, but rather
the complete protein occupancy profiles, which are best visualised 
experiment-wide on a PCA plot. As such, we prefer to focus on the
direct, qualitative comparison of individual replicate PCA plots,
assuring that each displays acceptable sub-cellular resolution. Note
that in the code chunk below, we mirror the x-axis to represent the
two figures with the same orientation. The interactive "compare"
application part of the \Biocpkg{pRolocGUI} package is also useful for
examining replicate experiments (see the next section interactive
visualisation for details).

@@

The above analysis is computationally intensive and best parallelised over
multiple workers. This phenoDisco analysis took ~24 hours to complete
when parallelised over 40 workers. As such, in the interest of time 
users can access the above results which are
pre-computed and stored along with the \texttt{pRolocdata} package. 
Please see the appendix to load these results.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{hl}
\end{alltt}
\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 5032 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126 X127N ... X131.1 (20 total)
##   varLabels: Replicate Tag Gradient.Fraction Iodixonal.Density
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
##   fvarLabels: uniprot.accession uniprot.id ... pd (27 total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Added markers from  'mrk' marker vector. Wed Apr 11 14:12:52 2018 
## Added markers from  'pdres' marker vector. Wed Apr 11 14:12:53 2018 
##  MSnbase version: 2.5.10
\end{verbatim}
\begin{alltt}
\hlkwd{getMarkers}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"pd"}\hlstd{)}
\end{alltt}
\begin{verbatim}
## organelleMarkers
##                          40S Ribosome 
##                                   106 
##                          60S Ribosome 
##                                    95 
## Endoplasmic reticulum/Golgi apparatus 
##                                   393 
##                         Mitochondrion 
##                                   525 
##                           Phenotype 1 
##                                   300 
##                           Phenotype 2 
##                                   253 
##                           Phenotype 3 
##                                   203 
##                           Phenotype 4 
##                                    74 
##                           Phenotype 5 
##                                    91 
##                       Plasma membrane 
##                                   421 
##                            Proteasome 
##                                    92 
##                               unknown 
##                                  2479
\end{verbatim}
\end{kframe}
\end{knitrout}

We can plot the results using the \texttt{plot2D} function (Figure
\ref{fig:plotPDres}).

@@

In the code chunk below we then pass these weights 
to the \texttt{svmOptimisation} function. Once again, we provide
the optimisation results for users to load directly if they wish to save
computational time, please see the appendix to load these results.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{hl} \hlkwb{<-} \hlkwd{svmClassification}\hlstd{(hl, params,} \hlkwc{class.weights} \hlstd{= w,} \hlkwc{fcol} \hlstd{=} \hlstr{"markers"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

There are many ways to set thresholds and the choice of method will
depend on the biological question and experimental design at hand. One
viable approach in the frame of the above experimetal design would be
to manually set a FDR, say 5\%, per organelle. To do this the user
would examine the top scoring predictions for each organelle, and then
set a threshold at the score at which they achieve 5\% of false
assignments per organelle.  The definintion of a false assignment
would depend on the information available, for example, validity or
lack of validity for the localisation from another experiment as
reported in the literature or a reliable database.  If such
information is not available, one crude method is to set a threshold
per organelle by extracting the median or 3rd quartile score per
organelle.  For example, in the code chunk below, we use the
\texttt{orgQuants} function to extract the median organelle scores and
then pass these scores to the \texttt{getPredictions} function to
extract the new localisations that meet this scoring criteria. Any
sub-cellular predictions that fall below the specified thresholds are
labelled as unknown.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{(ts} \hlkwb{<-} \hlkwd{orgQuants}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{=} \hlnum{.5}\hlstd{))}
\end{alltt}
\begin{verbatim}
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##               0.4336250               0.3041097               0.3895958 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##               0.6855842               0.5981495               0.4333482 
##    Extracellular matrix                Lysosome           Mitochondrion 
##               0.4243711               0.5982753               0.9494273 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##               0.7951711               0.7080911               0.3141263 
##         Plasma membrane              Proteasome 
##               0.7170268               0.4188616
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##               0.4336250               0.3041097               0.3895958 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##               0.6855842               0.5981495               0.4333482 
##    Extracellular matrix                Lysosome           Mitochondrion 
##               0.4243711               0.5982753               0.9494273 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##               0.7951711               0.7080911               0.3141263 
##         Plasma membrane              Proteasome 
##               0.7170268               0.4188616
\end{verbatim}
\begin{alltt}
\hlstd{hl} \hlkwb{<-} \hlkwd{getPredictions}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{= ts)}
\end{alltt}
\begin{verbatim}
## ans
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##                      85                     171                      89 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##                     294                     474                     107 
##    Extracellular matrix                Lysosome           Mitochondrion 
##                      26                     124                     523 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##                     230                     344                      38 
##         Plasma membrane              Proteasome                 unknown 
##                     310                     159                    2058
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

We generally find supervised learning more suited to the task of
protein localisation prediction in which we use high-quality curated
marker proteins to build a classifier, instead of using an entirely
unsupervised approach to look for clusters and then look for
enrichment of organelles and complexes. In the latter we do not make
good use of valuable prior knowledge, and in our experience
unsupervised clustering can be extremely difficult due to
(i) the loose definition of what constitutes a cluster (for example 
whether it is defined by the quantitative data or the localisation 
information), (ii) the influence of the algorithm assumption on the 
cluster identification (for example parametric or non-parametric) 
and (iii) poor estimates of the number of clusters that may appear 
in the data.

@@

##  [1] gridExtra_2.3        pRolocdata_1.17.4    pRoloc_1.19.3       
